---
title: 'Population analysis'
output: github_document
---


```{r setup, include=FALSE, echo = FALSE}
# Basic knitr options
library(knitr)
opts_chunk$set(comment = NA, 
               echo = TRUE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.path = 'figures/')
```

Idenitifying the population living outside of a 5km radius of the nearest HF per district

### Libraries

```{r}
require(raster)
require(sp)
require(rgdal)
require(rts)
require(dplyr)
require(tidyr)
require(rgdal)
require(rgeos)
require(spatstat)
require(maptools)
require(gdata)
require(geosphere)
library(sf)
library(gdalUtils)
library(gridsample)
```

### Preparatory set-up

Populate your `data` folder with the following files:

```{r, echo = FALSE}
dir('data')
```

### Import shapefile

```{r}
cmrshp <- readOGR("data", "CMR_HDs")
```

### Import HF GPs coordinates

```{r}
hfpub <- read.csv("data/CMR_HFs_GPS_Public.csv")
hfpub <- na.omit(hfpub)
coordinates(hfpub) <- ~Long+Lat
proj4string(hfpub) <- proj4string(cmrshp)
```

### Import raster

```{r}
# Read the raster
# poprast <-raster("data/cmr_bsgme_v0a_100m_2018.tif")
poprast <- raster('data/cmr_ppp_2018.tif')#https://www.worldpop.org/geodata/summary?id=5849
# strata <- rasterize(cmrshp,poprast, field="Map_Code")
```

### Estimate the total population according to the entire file 

```{r}
sum(values(poprast), na.rm = T)
```

### Estimate the population per polygon

```{r}
# The below operation is slow, so saving a temp file to speed up on multiple runs
if('x.RData' %in% dir()){
  load('x.RData')
} else {
  x <- raster::extract(poprast, cmrshp, method = 'sum', na.rm = TRUE)
  save(x, file = 'x.RData')
}
populations <- lapply(x, sum, na.rm = T)
populations <- unlist(populations)
# Test to ensure we're on the right track
sum(populations)

# Insert the polygon-specific populations
# into the spatial df
cmrshp@data$population <- populations
```

## Create 5km buffers around health posts

```{r}
# devtools::install_github("valentinitnelav/geobuffer")
library(geobuffer)

# Get buffered areas for 5 km
km <- 5
bufs <- geobuffer_pts(xy = hfpub, dist_m = 100*km^3)
# bufs <- spTransform(bufs, CRS(cmrshp))
proj4string(bufs) <- proj4string(cmrshp)
```

## Get the polygon (district) for each health post

```{r}
# index numbers
indices <- sp::over(hfpub, polygons(cmrshp))

# names
districts <- cmrshp@data$District[indices]

# insert names and index numbers into dataset
hfpub@data$District <- districts
hfpub@data$District_index <- indices
```

## Get the population of the district of each health facility

```{r}
hfpub@data <- left_join(hfpub@data,
               cmrshp@data %>%
                 dplyr::select(District, population))
```


## Get the population within 5 km

```{r, eval = FALSE}
if('hfpub.RData' %in% dir()){
  load('hfpub.RData')
} else {
  hfpub@data$population_in_buffer <- NA
  hfpub@data$population_in_buffer_within_district <- NA
  for(i in 1:length(bufs)){
    this_buf <- bufs[i]
    this_district <- hfpub@data$District[i]
    this_shp <- cmrshp[cmrshp@data$District == this_district,]
    this_buf_in_district <- rgeos::gIntersection(this_buf, this_shp)
    pop_in_buf <- sum(unlist(raster::extract(poprast, this_buf, method = 'sum', na.rm = TRUE)), na.rm = TRUE)
    pop_in_buf_and_district <- sum(unlist(raster::extract(poprast, this_buf_in_district, method = 'sum', na.rm = TRUE)), na.rm = TRUE)
    # Put into the hf dataframe
    hfpub@data$population_in_buffer[i] <- pop_in_buf
    hfpub@data$population_in_buffer_within_district[i] <- pop_in_buf_and_district
  message(i)
  }
  save(hfpub, file = 'hfpub.RData')
}
```

## Simple arithmetic

Calculate amount of people within district but outside of buffer area

```{r, eval = FALSE}
hfpub@data$magic_number <-
  hfpub@data$population -
  hfpub@data$population_in_buffer_within_district
```


## Create dataset

Create a data base with one row per pixel and its respective population value 

```{r}
xyz <- rasterToPoints(poprast)
xyz <- data.frame(xyz)
```


